\documentclass[12pt]{exam}
\usepackage{amsmath,amsthm,amsbsy,amsfonts}
 
\boxedpoints
\pointsinmargin
\addpoints
 
\newcommand{\Z}{\mathbb Z}
\newcommand{\E}{\mathbb E}
\newcommand{\R}{\mathbb R}
\newcommand{\Q}{\mathbb Q}
 
\pagestyle{headandfoot}
%\extraheadheight{.4in}
%\extrafootheight{-1in}
%\extrawidth{.7in}
\firstpageheader{Calculus II}{\bf\Large Worksheet}{}
 
\begin{document}
\textbf{\hspace{-.3in}
Work = Force * Distance \\
Hydrostatic Force = Density * Depth * Area\\
Mass = Density * Area\\
Moment = Distance * Mass\\
Center of Mass = $(\overline{x},\overline{y})$ = $\left(\frac{M_y}{m},
\frac{M_x}{m}\right)$
}

\begin{questions}

\question A spring has a natural length of 5 ft.  A force of
4 lb stretches this spring 1/4 ft.  
\begin{parts}
\part Find the work done in stretching this spring from its natural
length to 7 ft. 
\[
\int_0^2 16 x \, dx
\]

\part Find the work done in stretching this spring from 8 ft to 9 ft.
\[
\int_3^4 16x \, dx
\]
\end{parts}

\question How much work is done in emptying a cylindrical tank of 
radius 5 m and height 10 m by pumping the the water through a 
pipe extending 4 m above the tank.  (Hint: Water weighs 9800 N/$m^3$, or
62.4 lb/$ft^3$.)
\[
\int_0^{10} 9800 \pi(5)^2 (y+4) \, dy = 
\int_0^{10} 9800 \pi(5)^2 (14-y) \, dy =
\int_4^{14} 9800 \pi(5)^2 y \, dy
\]

\question A rectangular gate built into a dam is 4 ft high and 8 ft wide.
The top of the gate is 2 ft below the surface of the water.  What is the 
force on the gate?
\[
\int_0^4 62.4(8)(y+2) \,dy =
\int_0^4 62.4(8)(6-y)\, dy=
\int_2^6 62.4(8)(y) \,dy
\]

\question What would be  the force if the gate were the upper half of disc of
radius 4 (and the top of the gate is 2 ft below the surface of the water)?
\[
\int_0^4  62.4 (6-y)\sqrt{4^2 - y^2} \, dy
\]

\question Find the mass, moments, and center of mass of a the
triangular lamina bounded by the $x$- and $y$- axes and $y=4-2x$.
(As always, assume the density of this lamina to be $1$.)
\begin{align}\notag
m&=\int_0^2 4-2x\,dx 
=\int_0^4 \frac{1}{2}\left(\frac{4-y}{2}\right) \,dy \\\notag
M_y&=\int_0^2 x(1)(4-2x)\, dx
=\int_0^4 \frac{1}{2}\left(\frac{4-y}{2}\right)(1)
\left(\frac{4-y}{2}\right)\,dy \\\notag
M_x&=\int_0^2 \frac{1}{2}(4-2x) (1) (4-2x)\,dx 
=\int_0^4 y(1)\left(\frac{4-y}{2}\right)\,dy \\\notag
\end{align}



\end{questions}
\end{document}